Multiple quantum NMR in solids as a method of a determination of Wigner--Yanase skew information

### 2021-02-17 manuscript.tex
Fix bracket in math equation
Fix typos

\begin{abstract}
%% Text of abstract
A connection of the Wigner--Yanase skew information and multiple quantum (MQ) NMR coherences is considered at different temperatures and evolution times of nuclear spins with the dipole-dipole interactions in MQ NMR experiments in solids.
It is shown that the Wigner--Yanase skew information at temperature is equal to the doubled second moment of the MQ NMR spectra at the temperature twice large and any evolution times.
A comparison of the many--spin entanglement obtained with the Wigner--Yanase information and the Fisher information is conducted.
\end{abstract}

@@

\section{Introduction}
\label{sec:1}
The Wigner--Yanase skew information~\cite{1,2,3} together with the Fisher information~\cite{4,5} allow to develop the powerful methods for an investigation of entanglement, including the many--particle entanglement~\cite{6,7}.
Further investigations of the many--particle entanglement require a development of the corresponding experimental methods.
In particular, it was shown~\cite{6,8} that a low bound on the quantum Fisher information~\cite{4,5} coincides with the double second moment of the spectrum of multiple quantum (MQ) coherences.
As a result, the low bound on the quantum Fisher information can be found in MQ NMR experiments~\cite{9},
in cold--atom experiments, including Bose--Einstein condensates, ultracold atoms in cavities, and trapped ions~\cite{10,11,12,13,14}.
Using the properties of the quantum Fisher Information one can obtain a number of entangled particles (spins) in the system under consideration~\cite{6} and even to find the dependence of a number of entangled spins on temperature~\cite{8}.

The Wigner--Yanase skew information~\cite{1,2,3} is also connected with the spectrum of MQ coherences.
In particular, we demonstrate in the present article that the Wigner--Yanase skew information in a spin system $(S = 1/2)$ with the dipole--dipole interactions (DDI) in MQ NMR experiment~\cite{9} at the system temperature $T$ equals to the double second moment of the MQ NMR spectrum obtained at temperature $2T$.
Using the properties of the Wigner--Yanase skew information one can investigate the many--spin entanglement on the basis of the MQ NMR spectroscopy~\cite{9}.

@@

MQ NMR methods are widely used for solving problems of quantum informatics~\cite{16,17}. The MQ NMR experiment consists of four distinct periods of time Fig.~(\ref{fig:1}): preparation $(\tau)$, evolution $(t_1)$, mixing $(\tau)$, and detection $(t_2)$~\cite{9}.
MQ NMR coherences are created by a periodic multipulse sequence, consisting of $I_x$-pulses, irradiating the system of the preparation period~\cite{9}.
If the inverse period of the multipulse sequence significantly exceeds the local dipolar field (in frequency units)~\cite{18} 
MQ NMR dynamics can be described by the averaged nonsecular two--spin/two--quantum Hamiltonian~$H_{MQ}$
%
\begin{equation} \label{eq:1}
        H_{MQ} = H^{(+2)} + H^{(-2)} , \quad
        H^{(\pm 2)} = -\frac{1}{2} \sum_{j<k} D_{jk} I_{j} ^\pm I_k^\pm,
\end{equation}
%
where $D_{jk}$ is the coupling constant between spins $j$ and $k$, and $I_{j} ^+ I_k ^-$ are the raising and lowering operators of spin $j$.
On the mixing period the spin system is irradiated by the multiple sequence with $\pm y$--pulses.
As a result, the averaged nonsecular two-spin/two quantum Hamiltonian on the mixing period equals $(-H_{MQ})$~\cite{9}.

@@

\begin{equation}
    \label{eq:2}
        i\frac{d\rho(t)}{dt} = [H_{MQ}, \rho(t)]
\end{equation}
with the initial thermodynamic equilibrium density matrix
\begin{equation}
    \label{eq:3}
        \rho(0) = \rho_\mathrm{eq} = \frac{\exp(\frac{\hbar \omega_0}{kT}I_z)}{Z},
\end{equation}
where $Z=Tr \left\{e\frac{\hbar \omega_0}{kT}I_z \right\}$ is the partition function, $\hbar$ and $k$ are the Pauli and Boltzmann constants, $\omega_0$ is the Larmor frequency, $T$ is the temperature, and $I_z$ is the operator of the projection of the total spin angular momentum on the $z$-axis, which is directed along the strong external magnetic field.

@@

\section{The Wigner--Yanase skew information and MQ NMR}
\label{sec:3}
According definition~\cite{1,2} the Wigner--Yanase skew information
\begin{equation}
    \label{eq:14}
        I_{WY}(\rho(\tau,\beta),I_z) = -\frac{1}{2}
            Tr([\sqrt{\rho(\tau,\beta)},\sigma_z])^2 =
                -2Tr([\sqrt{\rho(\tau,\beta)},I_z])^2,
\end{equation}
where the Pauli operator $\sigma_z=2I_z$.
Introducing the evolution operator
\begin{equation}
    \label{eq:15}
        V(\tau) = e^{iH_{MQ}\tau}
\end{equation}
and using Eq.~(\ref{eq:3}) one can write the density matrix $\rho(\tau,\beta)$ as follows:
\begin{equation}
    \label{eq:16}
        \rho(\tau,\beta) = V^+(\tau) \frac{e^{\beta I_z}}{Z}V(\tau)
\end{equation}
Now we use te evident relationship:
\begin{equation}
    \label{eq:17}
        \sqrt{\rho(\tau,\beta)} =
            \sqrt{V^+(\tau)\frac{e^{\beta I_z}}{Z}V(\tau)} =
                V^+(\tau) \frac{e^{\frac{\beta}{2}}I_z}{\sqrt{Z}}V(\tau).
\end{equation}
it can be proved by simple calculation:
\begin{equation}
    \label{eq:18}
        \sqrt{\rho}\sqrt{\rho} =
            V^+(\tau)\frac{e^\frac{\beta}{2}I_z}{\sqrt{Z}}
                V(\tau)V^+(\tau)\frac{e^{\frac{\beta}{2}}I_z}{\sqrt{Z}}V(\tau) =
            V^+(\tau)\frac{e^\beta I_z}{Z}V(\tau) =
        \rho(\tau,\beta)
\end{equation}
%
Then we have
%
\begin{equation} \label{eq:19}
    \left[I_z,\sqrt{\rho(\tau,\beta)}\right] 
    = \left[I_z, \sum_k \rho_k \left(\tau, \frac{\beta}{2}\right)\right] 
    = \sum_k k\rho_k \left(\tau, \frac{\beta}{2}\right),
\end{equation}
%
and
%
\begin{equation} \label{eq:20}
    Tr[I_z,\sqrt{\rho(\tau,\beta)}]^2 
    = Tr\left\{\sum_{k,k'}kk'
        \rho_k\left(\tau,\frac{\beta}{2}\right)
        \rho_k\left(\tau,\frac{\beta}{2}\right) 
    \right\}
    = \sum_k k^2 J_k\left(\tau,\frac{\beta}{2}\right).
\end{equation}
%
Finally, one can obtain that
%
\begin{equation} \label{eq:21}
    I_{WY}\left(\rho(\tau, \beta), I_z\right)
    = 2\sum_k k^2 J_k\left(\tau, \frac{\beta}{2}\right)
    = 2M_2\left(\tau, \frac{\beta}{2}\right)
\end{equation}
%
Thus, we obtain the important assertion.
If the spin system is investigated with MQ NMR at temperature $T\sim\beta^{-1}$ then the Wigner--Yanase skew information equals to the doubled second momentum of the distribution of the intensities of the MQ NMR coherences at temperature $2T \sim 2\beta^{-1}$ at any moment of the spin evolution.

The Wagner-Yanase skew information is connected with the second moment of the distribution of the MQ NMR coherences analogously to the Fisher information.
We compare these information the following Section~\ref{sec:4}.

@@

\begin{figure}
	\includegraphics[width=0.95\linewidth]{entangled_spins_by_temp}
	\caption{
		The dependence of the number of the entangled spins on the inverse temperature $\beta = \frac{\pi \omega_0}{kT}$;
		black circles - the results are obtained with the Fisher information;
		open circles - results are obtained with the Wigner--Yanase information.
	}
	\label{fig:2}
\end{figure}

The Wigner--Yanase skew information $I_{WY}(\rho(\tau,\beta),I_z)$ and the Fisher information $I_F(\rho(\tau,\beta),I_z)$ can be used for an investigation of the many-spin entanglement.
The informations are connected by the following restriction~\cite{3}
%
\begin{equation} \label{eq:22}
    I_{WY}\left(\rho(\tau,\beta), I_z\right)
    \leq I_F\left(\rho(\tau,\beta), I_z\right)
    \leq 2I_{WY}\left(\rho(\tau,\beta), I_z\right).
\end{equation}
%
The restrictions (22) allow us to hope that the obtained results of the number of entangled spins are not strong by different.
For the comparison we used the model~\cite{21} of a nonspherical nanopore filled with a gas of spin-carrying atoms (for example, xenon) or molecules in a strong external magnetic field.
This model allows to investigate the many-spin entanglement in the spin system consisting of hundreds of nuclear spins~\cite{7}.

>>>>

\begin{abstract}
%% Text of abstract
A connection of the Wigner--Yanase skew information and multiple quantum (MQ) NMR coherences is considered at different temperatures and evolution times of nuclear spins with the dipole-dipole interactions in MQ NMR experiments in solids.
It is shown that the Wigner--Yanase skew information at temperature $T$ is equal to the doubled second moment of the MQ NMR spectra at the temperature twice large and any evolution times.
A comparison of the many--spin entanglement obtained with the Wigner--Yanase information and the Fisher information is conducted.
\end{abstract}

@@

\section{Introduction}
\label{sec:1}
The Wigner--Yanase skew information~\cite{1,2,3} together with the Fisher information~\cite{4,5} allow to develop the powerful methods for an investigation of entanglement, including the many--particle entanglement~\cite{6,7}.
Further investigations of the many--particle entanglement require a development of the corresponding experimental methods.
In particular, it was shown~\cite{6,8} that a low bound on the quantum Fisher information~\cite{4,5} coincides with the double second moment of the spectrum of multiple quantum (MQ) coherences.
As a result, the low bound on the quantum Fisher information can be found in MQ NMR experiments~\cite{9},
in cold--atom experiments, including Bose--Einstein condensates, ultracold atoms in cavities, and trapped ions~\cite{10,11,12,13,14}.
Using the properties of the quantum Fisher information one can obtain a number of entangled particles (spins) in the system under consideration~\cite{6} and even to find the dependence of a number of entangled spins on temperature~\cite{8}.

The Wigner--Yanase skew information~\cite{1,2,3} is also connected with the spectrum of MQ coherences.
In particular, we demonstrate in the present article that the Wigner--Yanase skew information in a spin system $(s = 1/2)$ with the dipole--dipole interactions (DDI) in MQ NMR experiment~\cite{9} at the system temperature $T$ equals to the double second moment of the MQ NMR spectrum obtained at temperature $2T$.
Using the properties of the Wigner--Yanase skew information one can investigate the many--spin entanglement on the basis of the MQ NMR spectroscopy~\cite{9}.

@@

MQ NMR methods are widely used for solving problems of quantum informatics~\cite{16,17}. The MQ NMR experiment consists of four distinct periods of time Fig.~(\ref{fig:1}): preparation $(\tau)$, evolution $(t_1)$, mixing $(\tau)$, and detection $(t_2)$~\cite{9}.
MQ NMR coherences are created by a periodic multipulse sequence, consisting of
$\pm$-pulses, irradiating the system of the preparation period~\cite{9}.
If the inverse period of the multipulse sequence significantly exceeds the local dipolar field (in frequency units)~\cite{18}
MQ NMR dynamics can be described by the averaged nonsecular two--spin/two--quantum Hamiltonian~$H_{MQ}$
%
\begin{equation} \label{eq:1}
        H_{MQ} = H^{(+2)} + H^{(-2)} , \quad
        H^{(\pm 2)} = -\frac{1}{2} \sum_{j<k} D_{jk} I_{j} ^\pm I_k^\pm,
\end{equation}
%
where $D_{jk}$ is the coupling constant between spins $j$ and $k$, and $I_{j} ^+ I_k ^-$ are the raising and lowering operators of spin $j$.
On the mixing period the spin system is irradiated by the multiple sequence with $\pm y$--pulses.
As a result, the averaged nonsecular two-spin/two quantum Hamiltonian on the mixing period equals $(-H_{MQ})$~\cite{9}.

@@

\begin{equation}
    \label{eq:2}
        i\frac{d\rho(t)}{dt} = [H_{MQ}, \rho(t)]
\end{equation}
with the initial thermodynamic equilibrium density matrix
\begin{equation}
    \label{eq:3}
        \rho(0) = \rho_\mathrm{eq} = \frac{\exp(\frac{\hbar \omega_0}{kT}I_z)}{Z},
\end{equation}
where $Z=Tr \left\{exp\left(\frac{\hbar \omega_0}{kT}I_z\right) \right\}$ is the partition function, $\hbar$ and $k$ are the Pauli and Boltzmann constants, $\omega_0$ is the Larmor frequency, $T$ is the temperature, and $I_z$ is the operator of the projection of the total spin angular momentum on the $z$-axis, which is directed along the strong external magnetic field.

@@

\section{The Wigner--Yanase skew information and MQ NMR}
\label{sec:3}
According definition~\cite{1,2} the Wigner--Yanase skew information
\begin{equation}
    \label{eq:14}
        I_{WY}(\rho(\tau,\beta),I_z) = -\frac{1}{2}
            Tr([\sqrt{\rho(\tau,\beta)},\sigma_z])^2 =
                -2Tr([\sqrt{\rho(\tau,\beta)},I_z])^2,
\end{equation}
where the Pauli operator $\sigma_z=2I_z$.
Introducing the evolution operator
\begin{equation}
    \label{eq:15}
        V(\tau) = e^{iH_{MQ}\tau}
\end{equation}
and using Eq.~(\ref{eq:3}) one can write the density matrix $\rho(\tau,\beta)$ as follows:
\begin{equation}
    \label{eq:16}
        \rho(\tau,\beta) = V^+(\tau) \frac{e^{\beta I_z}}{Z}V(\tau)
\end{equation}
Now we use te evident relationship:
\begin{equation}
    \label{eq:17}
        \sqrt{\rho(\tau,\beta)} =
            \sqrt{V^+(\tau)\frac{e^{\beta I_z}}{Z}V(\tau)} =
                V^+(\tau) \frac{e^{\frac{\beta}{2}}I_z}{\sqrt{Z}}V(\tau).
\end{equation}
It can be proved by simple calculation:
\begin{equation}
    \label{eq:18}
        \sqrt{\rho}\sqrt{\rho} =
            V^+(\tau)\frac{e^\frac{\beta}{2}I_z}{\sqrt{Z}}
                V(\tau)V^+(\tau)\frac{e^{\frac{\beta}{2}}I_z}{\sqrt{Z}}V(\tau) =
            V^+(\tau)\frac{e^\beta I_z}{Z}V(\tau) =
        \rho(\tau,\beta)
\end{equation}
%
Then we have
%
\begin{equation} \label{eq:19}
    \left[I_z,\sqrt{\rho(\tau,\beta)}\right]
    = \left[I_z, \sum_k \rho_k \left(\tau, \frac{\beta}{2}\right)\right]
    = \sum_k k\rho_k \left(\tau, \frac{\beta}{2}\right),
\end{equation}
%
and
%
\begin{equation} \label{eq:20}
	Tr\left[I_z,\sqrt{\rho(\tau,\beta)} \right]^2
	= Tr\left\{\sum_{k,k'}kk'
		\rho_k\left(\tau,\frac{\beta}{2}\right)
		\rho_k'\left(\tau,\frac{\beta}{2}\right)
	\right\}
	= \sum_k k^2 J_k\left(\tau,\frac{\beta}{2}\right).
\end{equation}
%
Finally, one can obtain that
%
\begin{equation} \label{eq:21}
    I_{WY}\left(\rho(\tau, \beta), I_z\right)
    = 2\sum_k k^2 J_k\left(\tau, \frac{\beta}{2}\right)
    = 2M_2\left(\tau, \frac{\beta}{2}\right)
\end{equation}
%
Thus, we obtain the important assertion.
If the spin system is investigated with MQ NMR at temperature $T\sim\beta^{-1}$ then the Wigner--Yanase skew information equals to the double second momentum of the distribution of the intensities of the MQ NMR coherences at temperature $2T \sim 2\beta^{-1}$ at any moment of the spin evolution.

The Wagner-Yanase skew information is connected with the second moment of the distribution of the MQ NMR coherences analogously to the Fisher information.
We compare these informations the following Section~\ref{sec:4}.

@@

\begin{figure}
	\includegraphics[width=0.95\linewidth]{entangled_spins_by_temp}
	\caption{
		The dependence of the number of the entangled spins on the inverse temperature $\beta = \frac{\pi \omega_0}{kT}$;
		black circles - the results are obtained with the Fisher information;
		open circles - the results are obtained with the Wigner--Yanase information.
	}
	\label{fig:2}
\end{figure}

The Wigner--Yanase skew information $I_{WY}(\rho(\tau,\beta),I_z)$ and the Fisher information $I_F(\rho(\tau,\beta),I_z)$ can be used for an investigation of the many-spin entanglement.
The informations are connected by the following restriction~\cite{3}
%
\begin{equation} \label{eq:22}
    I_{WY}\left(\rho(\tau,\beta), I_z\right)
    \leq I_F\left(\rho(\tau,\beta), I_z\right)
    \leq 2I_{WY}\left(\rho(\tau,\beta), I_z\right).
\end{equation}
%
For the comparison we used the model~\cite{21} of a nonspherical nanopore filled with a gas of spin-carrying atoms (for example, xenon) or molecules in a strong external magnetic field.
The restrictions (22) allow us to hope that the obtained results of the number of entangled spins are not strongly different.
This model allows to investigate the many-spin entanglement in the spin system consisting of hundreds of nuclear spins~\cite{7}.
